MS_Conclusion

### 2023-04-19 Markdown/Manuscript/Child_RMDs/Clippings.Rmd
first pass and updated bib

---
title: "MS_Discussion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

>>>>

---
title: "Clippings"
output: html_document
date: "2023-04-18"
---

#################################

Computational models, particularly those addressing economic topics, have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, health economic models typically have a number of limitations that at best restrict their usefulness and which in some cases may facilitate harmful misuse.

There is a strong case for making health economists more accountable for the appropriate use and social acceptability of their models. Potential users of models should be able to assess their adequacy for a particular purpose [@thompson2019escape].  This goal is difficult to achieve in the current context of the poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017] and frequently insufficient validation [@Ghabri2019] of health economic models. Many health economic models are released with undeclared errors [@Radeva2020]. The value judgments that strongly shape health economic analyses are rarely made explicit, omissions that may lead to socially unacceptable policy recommendations [@duckett2022journey]. A modelling team's value judgments about what questions to address, the most important features of a system to represent and the weighting of different types of evidence may be poorly aligned with those of the people impacted by decisions informed by model analyses [@thompson2022escape]. As health economic models adopt more sophisticated techniques, the need for accountability grows. More complex models may be more prone to propagation errors [@Saltelli2019] and models designed to address multiple questions should be expected to meet more onerous verification and validation obligations [@Eddy2012; @Feenstra2022]. The nature and extent of individual model authorship contributions may be less clear in models implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape].

The resource and skills intensive undertaking of modelling could be made more tractable if it was easier for health economists to re-use each other's models [@Arnold2010]. However, as many health economic models are owned by pharmaceutical companies and consultancies, commercial considerations can limit the reuse of models and their constituent code and data [@Feenstra2022]. Legal and ethical issues such as privacy and confidentiality requirements further limit public release of some model artefacts [@Pouwels2022]. Transferring a health economic model developed for one jurisdiction for application in another typically involves retaining some features and updating others [@barbieri2010international]. However, models do not always make implementation choices (for example, use of concepts with standardised meanings across jurisdictions<!-- [@garcia2021cost] -->, supplying context specific data as replaceable data-packs and distributing source code under licenses that allow derivative works) that facilitate model transferability.

Health economic models should be updated and refined as new evidence emerges and decision contexts change [@Jenkins2021], but this occurs infrequently [@Sampson2017]. Funding for health economic modelling projects rarely extend to provision of medium term support for model updates and improvements. The career trajectories of health economists can also mitigate against adequate maintenance of a model, it being relatively common for model authors to have moved on from the team that owns the model and / or from working on the health condition for which the model was developed.

A potential strategy for improving model accountability, reusability and updatability is to make models both modular and open-source.

A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular]. Modular approaches are currently being used to facilitate the development of complex computational models in disciplines such as biology [@pan2021modular], ecology [@barros2023empowering] and neuroscience [@frazier2022robust]. In health economics the related and enabling concept of reference models has been recommended [@Afzali2013], but peer reviewed studies describing modular health economic models remain relatively rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013].

Modular models provide an opportunity for multiple modelling teams to contribute to, test and reuse models. To enhance this capacity, modular models may be implemented as open source projects that give others liberal permissions to access and use model source code and data [@pan2021modular; @barros2023empowering; @frazier2022robust]. Although there appears to be in principle support from many health economists for greater use of open source health economic models (OSHEMs)  [@Pouwels2022], actual implementations are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption of OSHEMs include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020], but guidelines for implementing OSHEMs remain scarce, piecemeal and need improving [@Sampson2019].

Our interest in modular and open source approaches developed when we began seeking am appropriate framework for undertaking and validly synthesising diverse types of economic research in mental health. Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge [@RN22;@RN23]. A substantial economic literature already exists to assess the affordability and value for money of mental health interventions [@RN34]. This economic evaluation work is an essential prerequisite for improving allocative efficiency in mental health, but could be of greater value to systems planners if integrated with a broader program of economic research.

We are developing ready4 (https://www.ready4-dev.com), a modular OSHEM (MOSHEM) on the mental health of young people aged 12 to 25. We are initially applying ready4 to four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:
<!-- There is therefore an opportunity for health economists to undertake mental health research of significant potential public benefit. -->

- *health and its value* (our projects: utility mapping models);

- *determinants of health and ill-health* (our projects: models for creating synthetic household populations with key risk and protective factors for mental disorders);

- *demand for health and health care* (our projects: spatial epidemiology and help-seeking choice models); and

- *supply of health services* (our projects: a model of primary mental health care services).

Once these projects are completed, our aim is to flexibly combine these models to answer questions in two additional Wagstaff and Culyer domains:

- *efficiency and equity* (our goal: assess the distributional impacts and identify the optimal targeting of care provision); and

- *economic evaluation* (our goal: assess the cost-utility of competing policy options for improving the mental health of young people).

Standardisation is an essential enabler of automation, collaboration, interoperability and transferability in modular and open source approaches. An early requirement to implement ready4 is therefore to identify an appropriate set of standards. However, we are not aware of any consolidated source of guidance about recommended standards for MOSHEMs. Once identified, the standards specified for ready4 will need to be implemented by software, some of which will require custom development.

<!-- When planning how to undertake and synthesise this program of work, we became concerned with issues relating to accountability, reusability and updatability. If mistakes are common in individual health economic modelling projects [@Radeva2020], then combining multiple projects may be particularly error prone. The risk of errors may also depend on whether a model is too simple or too complex [@Saltelli2019], so the level of granularity appropriate for one model application may be ill-suited to another. Undertaking a multiannual program of interdependent work raises the risk that models developed in the first phase of a project may need to be updated by the time the project is approaching conclusion. -->

<!-- These concerns led us to seek the MOSHEM approach that provides the transparency to make it easier for errors to be uncovered, the adaptability to fix these errors and the flexibility to facilitate interchangeability between representations of different levels of granularity.  -->

### Standards for an accountable MOSHEM
Guidance on transparency in health economic modelling published over ten years ago [@Eddy2012] made recommendations on documenting models but notably did not include recommendations on sharing model code and data. However, more recent and multidisciplinary healthcare modelling guidance [@Erdemir2020] recommends using existing digital repository services to make these types of digital model artefacts publicly available. Some repositories such as GitHub [@github2007] provide tools for disseminating work in progress code and providing highly transparent records of the complete development history and individual authorship contributions of a software project. Other repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection.

Model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Consistent use of meaningful naming conventions when authoring code is recommended [@Wilson_2017; @Alarid2019]. Code can be made easier to follow by using the practices of abstraction [@8717448], where only simple, high level commands are routinely exposed to reviewers, and polymorphism [@7181447], where the same command (e.g. "simulate") can be reused to implement different algorithms of the same type. Programs to implement model analyses can be made comprehensible to even non-technical users through the use of literate programming techniques and tools like RMarkdown [@xie2018r] that integrate computer code with plain English descriptions.

An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. One useful concept for informing model users about the extensiveness of verification checks is code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. <!-- In R, the testthat [@testthat2011] and covr [@covr2020] tools can be used in conjunction with GitHub to define tests and report coverage metrics. -->Transcription errors - mistakes introduced when transferring data between sources, models and reports - are very common in health economic models [@Radeva2020]. The risk of these errors might be lower if there was full transparency across all steps in a study workflow. Scientific computing tools now make it relatively straightforward to author programs that reproducibly execute all steps in data ingest, processing and reporting [@Wilson_2017].

Code and data should be distributed with tools that make it easy for potential users to appropriately cite each model artefact.

### Standards for a reusable MOSHEM
```{r echo = F}
gpl_pc_dbl <- get_license_share("GPL")
```
To make model code and data widely re-usable by others, it is important to provide users with appropriate and explicit permissions. In the context of open source models, there are two broad categories of licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which the content could be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open source arrangements. For code, it may be appropriate to adopt the prevailing open source licensing practice within the programming language being used. For data, it may not be sufficient to simply choose between a permissive license like the Public Domain Dedication (CC0) [@cc02022] or a copyleft option such as the Attribution-Share Alike (CC-BY-SA) [@bysa2022]. In addition to ensuring that data is ethically appropriate for disseminate in open access repositories, responsible custodianship of some de-identified or aggregated data may involve using or adapting template terms of use [@sampleterms2022] which have a number of ethical clauses (for example, prohibiting efforts to re-identify research participants).

Storing model code and data in separate files and locations (as opposed to hard coding - embedding data into source code) can make it easier to apply models to different decision contexts and to selectively restrict access to data that are confidential, while disseminating all other model artefacts. Clear distinctions should be made between model modules (code that defines abstract data structures and the algorithms that can be applied to data described by these structures), model datasets (digital information such as parameter values, unit records, etc) and model analyses (code that links model datasets to model modules and specifies the algorithms to apply to data associated with each module).

The software development practice of encapsulation [@8717448] can be used to help ensure that model modules continue to work as intended when they are combined [@ready4oop2022]. In some cases, combining modules may mean new versions of modules have to be created to better account for interaction effects. The concept of inheritance [@8717448] can be used to write code that efficiently achieves this objective as well as to facilitate selective editing of modules when transferring models to different decision contexts [@ready4oop2022]. Writing algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. Functions to implement model algorithms can be associated with data structures (also known as a class) via a special type of function called a method. Model modules of a similar type or purpose can be efficiently distributed and documented by bundling them as code libraries. It is good practice to make available test or toy data to demonstrate the use of model algorithms [@Wilson_2017].

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them for out of sample prediction [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when publicly releasing model artefacts derived from data on human subjects as they may by default embed a copy of the source dataset. Sensitive dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users.

### Standards for an updatable MOSHEM
To avoid MOSHEMs going stale - losing validity and usefulness with time - they should be routinely updated.  Each update of code and data should be uniquely identifiable and retrievable, a goal that can be facilitated by use of version control tools [@Erdemir2020] such as git [@git20XX].

Potential users of model code and data should be able to easily identify the appropriateness of a version to their needs. Using semantic versioning [@semver20xx] conventions can signal the potential importance of an update to users of model code and data. Also informative is to clearly label the type of code versions hosted in a repository as either:

- "development" - typically the most comprehensive and up to date source code, but potentially not yet sufficiently well documented and tested for potential users to apply to purposes other than testing;

- "production" - code that has been released with a view to it being applied to its stated purposes and therefore typically required to meet defined documentation and testing standards; or

- "archive" - permanent copies of code at key milestones in its development that can aid study reproductions by making available the version of code used in a specific analysis.

Continuous integration [@CI2017] tools can help verify that each code update continues to pass a standardised battery of quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk of model revisions having unintended consequences on third party users.

We developed a toolkit to help us develop and use MOSHEM modules, datasets and analyses that meet the standards listed in Table \@ref(tab:timelygls). The toolkit is comprised of online repositories and software.

## Repositories <!-- Existing ref -->
We created a GitHub organisation (a collection of code repositories) where all our development code is stored and version controlled [@ready4gh2022]. We configured the repositories in our GitHub organisation to use GitHub actions to support continuous integration. Some of the continuous integration checks we have defined assess each library's compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022], to which we plan submitting future production releases. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022].

We enabled GitHub Pages in each repository we used for code library development to facilitate the creation and hosting of library documentation websites. We also developed a consolidated and versioned project documentation website [@rfwn2022] using the Hugo framework [@hugo_2023] and Docsy theme [@docsy_2023]. To host that website we established an account with Netlify [@netlify_2023] and linked that service to our GitHub organisation so that the website would automatically update whenever the source code in its GitHub repository was edited.

We also created a Zenodo community [@ready4zen2022] - a collection of permanent, uniquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we specify a version of code in one of our GitHub repositories as a "release", a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation [@ready4dv2022].

## Software <!-- Novel ref -->
We created six development version R libraries to help us author model modules, supply those modules with data and implement reproducible modelling analyses. The six libraries, their primary focus, the standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs).

A library called ready4 [@ready42022] defines a template module (using R's S4 class system) from which all model module data structures will inherit features and a novel syntax for attaching algorithms to those data structures. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs and for partially automating updates to the project documentation website.

Three R libraries are designed to standardise and partially automate workflows for authoring new model modules. The ready4pack library [@ready4pack2022] is designed to integrate with our GitHub organisation and provides tools for authoring module libraries that are:

- documented (with a website, a manual itemising selected contents and a manual itemising all contents);

- licensed (using GNU GPL-3 [@GNUGPL2022] by default);

- easily citable (citation information can be retrieved within an R session or from hosting repositories); and

- quality assured (each update triggers continuous integration workflows, including any unit tests created by module library authors).

The ready4pack library depends on two other module authoring libraries. Methods from the ready4fun library [@ready4fun2022] are used to verify that functions for implementing module algorithms are written in a consistent house style. That standardised format is then used by ready4fun methods to automatically generate basic documentation for each function. Methods from the ready4class [@ready4class2022] library are used to streamline and standardise the authoring of module data structures and the linking of methods to these data structures. Like ready4fun, the ready4class library uses standardised code implementation to automatically generate basic documentation for each module data structure.

The ready4use library  [@ready4use2022] contains modules for ingesting model datasets from online repositories (hosted on a Dataverse installation or on GitHub), labelling model datasets and sharing model datasets via online repositories. The ready4show library [@ready4show2022] contains tools to help author analyses programs that are either self-documenting or which trigger the creation of a scientific summary.

When used in conjunction with toolkit repositories, the six R libraries provide support for implementing 17 out of 20 framework standards (Table \@ref(tab:timelygls)). Standards relating to safe dissemination of statistical models (R8), user-interface development (R9) and deprecation conventions (U4) are better met through using existing third party R libraries. Preparing statistical models for dissemination can be accomplished with standard R data management tools like the dplyr [@dplyr2022] and purrr [@dplyr2022] libraries. User-interfaces are typically developed with the shiny [@shiny2022] library, for which a tutorial aimed at health economists is available [@SmithR2020]. The library lifecycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code.

We applied the framework to develop an initial set of ready4 modules, supply those modules with data and implement modelling analyses. These outputs were created as part of a previously described study [@Hamilton2021.07.07.21260129] to develop utility mapping models appropriate for use in samples of young people presenting to primary mental health services. The ready4 framework's modelling toolkit created the following artefacts:

 - development version module libraries for describing and validating youth mental health human record datasets [@hamilton_matthew_2022_6084467], scoring health utility [@hamilton_matthew_2022_6084824], specifying utility mapping models [@hamilton_matthew_2022_6116701] and implementing reproducible utility mapping studies [@gao_caroline_2022_6130155];

 - a development version library of functions for finding and using utility mapping models developed with these tools [@matthew_p_hamilton_2021_5646669];

 - data collections of synthetic populations for testing model modules [@DVN/HJXYKQ_2021] and study input and results data [@DVN/DKDIB0_2021];

 - programs for replicating all steps from data ingest to manuscript reporting [@hamilton_matthew_2022_6129906], applying utility mapping models to new data [@hamilton_matthew_2022_6416330] and generating a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821];

 - subroutines for creating a catalogue of utility mapping models [@hamilton_matthew_2022_6116385] and generating a draft scientific manuscript [@matthew_p_hamilton_2022_5976988] for studies implemented with these modules.

We created a checklist (Table \@ref(tab:checktb)) that we used to assess the extent to which study outputs met framework standards. We assessed these outputs as wholly or mostly meeting 18 out of 20 standards. The two standards where the study outputs currently fall short are in reporting code coverage and including a user-interface. Both these items are scheduled to be addressed when we release production versions of the code libraries.

### 2023-04-19 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
first pass and updated bib

In this paper, we describe:

(i) a framework that specifies MOSHEM standards and provides the tools to implement them; and

(ii) the application of that framework to develop and use an initial set of ready4 modules in a utility mapping study.

>>>>

In this paper, we describe:
(i)	a set of ethical responsibilities for CHEM developers and criteria for assessing responsible CHEM implementations;
(ii)	a prototype software framework for the ethical implementation of CHEMs; and
(iii)	use of the software framework to develop a computational economic model in youth mental health, with an initial focus on outcome valuation.

### 2023-04-19 Markdown/Manuscript/Child_RMDs/MS_Conclusion.Rmd
first pass and updated bib

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)`

>>>>

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)` 
We have identified criteria that can be used to systematically assess extent to which the computational implementation of health economic models adheres to the ethical goals of transparency, reusability and updatability. We have developed an open-source software framework that can support the ethical computational implementation of economic models in youth mental health. Our framework can be used as a prototype for developing future software frameworks to support ethical implementation of CHEMs.
